Inference of Gravitational Lens Time Delays Using Gaussian Processes

### 2013-05-03 doc/Leon/ms.tex
Covariance functions
Working on the document

% Document starts here
\begin{document}

Throughout this section we consider the measured signal in magnitudes,
rather than fluxes. Suppose that the underlying QSO variability as a function
of time is described by the following function:
\begin{eqnarray}
y_{\rm QSO}(t)
\end{eqnarray}
Then the noise-free light curves of the two images
$A$ and $B$ will be given by:
\begin{eqnarray}
y_A(t) &=& y_{\rm QSO}(t) + \mu_A(t)\\
y_B(t) &=& y_{\rm QSO}(t - \tau) + \mu_B(t).
\end{eqnarray}
The variability of image $A$ is the underlying QSO variability
plus a signal $\mu_A(t)$ that is due to microlensing or perhaps
other effects. The variability of image $B$ is given by
a lagged version of the QSO variability, plus another
perturbation $\mu_B(t)$ caused my microlensing.

The data are assumed to be noisy measurements of $y_A(t)$ and
$y_B(t)$ at various times. We denote the entire dataset (both images together)
as a vector of magnitude measurements:
\begin{eqnarray}
\yy = \{y_1, y_2, ..., y_N\}.
\end{eqnarray}

If we make the simplifying assumption that our prior knowledge about
$\mu_A(t)$, $\mu_B(t)$, and
$y_{\rm QSO}(t)$ can be described by Gaussian Processes \citep{rasmussen},
then the probability distribution for the data $\yy$ given the parameters
$\theta$
is a multivariate normal distribution, where the mean vector $\mm$
and the covariance matrix $\CC$ depend on the parameters:
\begin{eqnarray}
p(\yy | \theta) &=& \frac{1}{\sqrt{(2\pi)^N\det{\CC}}}
\exp\left[
-\frac{1}{2}
\left(
\yy - \mm
\right)^T
\CC^{-1}
\left(
\yy - \mm
\right)
\right]
\end{eqnarray}
The elements of the mean vector are just the mean magnitudes of the
appropriate images:
\begin{eqnarray}
m_i = 
\end{eqnarray}

@@

>>>>

% Document starts here
\begin{document}
In this section we describe the Bayesian model used to infer the
time delay from the data. This model will be described in greater detail in
a forthcoming contribution (Brewer and Marshall, in prep).
Throughout this section we consider the measured signal in magnitudes,
rather than fluxes, as the model assumptions (Section~\ref{sec:assumptions})
make more sense in terms of magnitudes.

\section{The Data}
We denote the entire dataset (both images together)
as a vector of magnitude measurements:
\begin{eqnarray}
\yy = \{y_1, y_2, ..., y_N\}.
\end{eqnarray}
Along with these measurements we have the timestamps $\{t_i\}_{i=1}^N$,
error bars $\{\sigma_i\}_{i=1}^N$ and flags
$\{c_i\}_{i=1}^N$ where a flag $c_i \in \{A, B\}$ tells us whether the
measurement is of image $A$ or image $B$. The likelihood will be
written as a probability distribution for $\yy$ given some parameters, and the
timestamps, errorbars and flags will be considered as part of the prior
information.

\section{Model Assumptions}\label{sec:assumptions}
Suppose that the underlying QSO variability as a function
of time, as seen through image $A$, is described by the following function:
\begin{eqnarray}
y_{\rm QSO}(t)
\end{eqnarray}
Then the light curves of the two images
$A$ and $B$ will be given by:
\begin{eqnarray}
y_A(t) &=& M_A + y_{\rm QSO}(t)\\
y_B(t) &=& M_B + y_{\rm QSO}(t - \tau).
\end{eqnarray}
The light curve of image $A$ is just the underlying QSO variability,
and the light curve of

plus a signal $\mu_A(t)$ that is due to microlensing or perhaps
other effects. The variability of image $B$ is given by
a lagged version of the QSO variability, plus another
perturbation $\mu_B(t)$ caused my microlensing.

The data are assumed to be noisy measurements of $y_A(t)$ and
$y_B(t)$ at various times.

If we make the simplifying assumption that our prior knowledge about
$\mu_A(t)$, $\mu_B(t)$, and
$y_{\rm QSO}(t)$ can be described by Gaussian Processes \citep{rasmussen},
then the probability distribution for the data $\yy$ given the parameters
$\theta$
is a multivariate normal distribution, where the mean vector $\mm$
and the covariance matrix $\CC$ depend on the parameters:
\begin{eqnarray}
p(\yy | \theta) &=& \frac{1}{\sqrt{(2\pi)^N\det{\CC}}}
\exp\left[
-\frac{1}{2}
\left(
\yy - \mm
\right)^T
\CC^{-1}
\left(
\yy - \mm
\right)
\right]
\end{eqnarray}
The elements of the mean vector are just the mean magnitudes of the
appropriate images:
\begin{eqnarray}
m_i = M_{c_i}
\end{eqnarray}

@@

We assume the following:
\begin{eqnarray}
\textnormal{Cov}\left(y_{\rm QSO}(t_1), y_{\rm QSO}(t_2)\right)
&=& \sigma_{\rm QSO}^2\exp\left[-\frac{\left|t_2 - t_1\right|}{L_{\rm QSO}}\right]
\end{eqnarray}

\begin{eqnarray}
\textnormal{Cov}\left(\mu_A(t_1), \mu_A(t_2)\right)
&=& \sigma_{{\rm ML}, A}^2\exp\left[-\left(\frac{\left|t_2 - t_1\right|}{L_{{\rm ML}, A}}\right)^\alpha\right]
\end{eqnarray}

\begin{eqnarray}
\textnormal{Cov}\left(\mu_B(t_1), \mu_B(t_2)\right)
&=& \sigma_{{\rm ML}, B}^2\exp\left[-\left(\frac{\left|t_2 - t_1\right|}{L_{{\rm ML}, B}}\right)^\alpha\right]
\end{eqnarray}
